Dominance Regions for Imaginary $\bfg$-vectors

### 2020-06-07 dominance_regions.tex
One direction mutations in rank two

  \section{Rank Two}
  Fix integers $b,c>0$ and assume $bc\ge4$.
  Let $B=\left[ \begin{array}{cc} 0 & -b \\ c & 0 \end{array} \right]$.
  Write $\cI \subset \RR^2$ for the \emph{imaginary cone} spanned by the vectors $(2b,-bc\pm\sqrt{bc(bc-4)})$.

@@

  \begin{theorem}
    For $\lambda\in\cI\cap\ZZ^2$, there are five classes of dominance polytopes.
    \begin{enumerate}
      \item if $\lambda$ lies in the interior of the cone spanned by the vectors $(2b,-bc-\sqrt{bc(bc-4)})$ and $(2,-c)$, then $\cC(\lambda)$ is the trapezoid with corner vertices $\lambda$, $(\frac{2-bc-\sqrt{bc(bc-4)}}{2}\lambda_1+\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_2,0)$, $(0,\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_1+\lambda_2)$, and ???.
      \item if $\lambda$ lies on the ray spanned by $(2,-c)$, say $\lambda=(2\ell,-c\ell)$, then $\cC(\lambda)$ is the triangle with corner vertices $\lambda$, $(2\ell-\frac{bc+\sqrt{bc(bc-4)}}{2}\ell,0)$, and $(0,\frac{\sqrt{bc(bc-4)}}{b}\ell)$.
      \item if $\lambda$ lies in the interior of the cone spanned by the vectors $(2,-c)$ and $(b,-2)$, then $\cC(\lambda)$ is the kite with corner vertices $\lambda$, $(\lambda_1+\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_2,0)$, $\frac{bc+\sqrt{bc(bc-4)}}{2\sqrt{bc(bc-4)}}(2\lambda_1+b\lambda_2,c\lambda_1+2\lambda_2)$, and $(0,\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_1+\lambda_2)$.
      \item if $\lambda$ lies on the ray spanned by $(b,-2)$, say $\lambda=(b\ell,-2\ell)$, then $\cC(\lambda)$ is the triangle with corner vertices $\lambda$, $(-\frac{\sqrt{bc(bc-4)}}{c}\ell,0)$, and $(0,\frac{bc+\sqrt{bc(bc-4)}}{2}\ell-2\ell)$.
      \item if $\lambda$ lies in the interior of the cone spanned by the vectors $(b,-2)$ and $(2b,-bc+\sqrt{bc(bc-4)})$, then $\cC(\lambda)$ is the triangle with corner vertices $\lambda$, $(\lambda_1+\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_2,0)$, $(0,\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_1+\frac{2-bc-\sqrt{bc(bc-4)}}{2}\lambda_2)$, and ???.
    \end{enumerate}
  \end{theorem}
  \begin{proof}
    Tedious calculation once we have proven what the slopes of the boundaries are.
  \end{proof}

>>>>

  \section{Rank Two}
  Fix integers $b,c>0$ and assume $bc\ge4$.
  Let $B=\left[ \begin{array}{cc} 0 & -b \\ c & 0 \end{array} \right]$.
  Write $\cI \subset \RR^2$ for the \emph{imaginary cone} spanned by the vectors $(2b,-bc\pm\sqrt{bc(bc-4)})$.
  \begin{lemma}
    The vectors $(2b, -bc+\sqrt{bc(bc-4)})$ (resp. $(2b, -bc-\sqrt{bc(bc-4)})$) and $(2c, bc+\sqrt{bc(bc-4)})$ (resp. $(2c, bc-\sqrt{bc(bc-4)})$) are perpendicular.
  \end{lemma}

@@

  \begin{lemma}
    \label{le:one direction}
    For $\lambda\in\cI\cap\ZZ^2$, 
    \begin{enumerate}
      \item $\cP_+(\lambda):=\bigcap_{i \ge 0}\phi_{t_i}^{-1}\cC(\phi_{t_i}\lambda)$ is the quadrilateral with corner vertices $\lambda$, $(\frac{2-bc-\sqrt{bc(bc-4)}}{2}\lambda_1-\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_2,0)$, $(0,\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_1+\lambda_2)$, and $(\frac{2-bc-\sqrt{bc(bc-4)}}{2}\lambda_1-b\lambda_2,\lambda_2)$.
      \item $\cP_-(\lambda):=\bigcap_{i \le 0}\phi_{t_i}^{-1}\cC(\phi_{t_i}\lambda)$ is the quadrilateral with corner vertices $\lambda$, $(\lambda_1+\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_2,0)$, $(0,-\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_1+\frac{2-bc-\sqrt{bc(bc-4)}}{2}\lambda_2)$, and $(\lambda_1,-c\lambda_1+\frac{2-bc-\sqrt{bc(bc-4)}}{2}\lambda_2)$.
    \end{enumerate}
  \end{lemma}

  \begin{theorem}
    For $\lambda\in\cI\cap\ZZ^2$, there are five classes of dominance polytopes.
    \begin{enumerate}
      \item if $\lambda$ lies in the interior of the cone spanned by the vectors $(2b,-bc-\sqrt{bc(bc-4)})$ and $(2,-c)$, then $\cC(\lambda)$ is the trapezoid with corner vertices $\lambda$, $(\frac{2-bc-\sqrt{bc(bc-4)}}{2}\lambda_1-\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_2,0)$, $(0,\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_1+\lambda_2)$, and ???.
      \item if $\lambda$ lies on the ray spanned by $(2,-c)$, say $\lambda=(2\ell,-c\ell)$, then $\cC(\lambda)$ is the triangle with corner vertices $\lambda$, $(2\ell-\frac{bc+\sqrt{bc(bc-4)}}{2}\ell,0)$, and $(0,\frac{\sqrt{bc(bc-4)}}{b}\ell)$.
      \item if $\lambda$ lies in the interior of the cone spanned by the vectors $(2,-c)$ and $(b,-2)$, then $\cC(\lambda)$ is the kite with corner vertices $\lambda$, $(\lambda_1+\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_2,0)$, $\frac{bc+\sqrt{bc(bc-4)}}{2\sqrt{bc(bc-4)}}(2\lambda_1+b\lambda_2,c\lambda_1+2\lambda_2)$, and $(0,\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_1+\lambda_2)$.
      \item if $\lambda$ lies on the ray spanned by $(b,-2)$, say $\lambda=(b\ell,-2\ell)$, then $\cC(\lambda)$ is the triangle with corner vertices $\lambda$, $(-\frac{\sqrt{bc(bc-4)}}{c}\ell,0)$, and $(0,\frac{bc+\sqrt{bc(bc-4)}}{2}\ell-2\ell)$.
      \item if $\lambda$ lies in the interior of the cone spanned by the vectors $(b,-2)$ and $(2b,-bc+\sqrt{bc(bc-4)})$, then $\cC(\lambda)$ is the triangle with corner vertices $\lambda$, $(\lambda_1+\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_2,0)$, $(0,-\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_1+\frac{2-bc-\sqrt{bc(bc-4)}}{2}\lambda_2)$, and ???.
    \end{enumerate}
  \end{theorem}
  \begin{proof}
    Following Lemma~\ref{le:one direction}, we see that the intersection $\cP_+(\lambda)\cap\cP_-(\lambda)$ degenerates to a triangle precisely when
    \[\frac{2-bc-\sqrt{bc(bc-4)}}{2}\lambda_1-\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_2=\lambda_1+\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_2\]
    or 
    \[\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_1+\lambda_2=-\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_1+\frac{2-bc-\sqrt{bc(bc-4)}}{2}\lambda_2.\]
    The first equations reduces to $c\lambda_1+2\lambda_2=0$, while the second reduces to $2\lambda_1+b\lambda_2=0$.
    In particular, the rays spanned by $(2,-c)$ and $(b,-2)$ determine this change of state in the dominance regions $\cP(\lambda)$.
  \end{proof}

  \begin{remark}
    Note that the rays which separate the regions inside $\cI$ correspond exactly to the columns of the associated Cartan matrix $\left[ \begin{array}{cc} 2 & -b \\ -c & 2 \end{array} \right]$.
    This unexpected coincidence is our reason for writing these results.
  \end{remark}
